Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy

### 2021-04-17 manuscript/manuscript.Rmd
More Suparna edits

---
title: "Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy"
short-title: "Who Cares About Crackdowns?"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
date: "October 2, 2020"
published: Working paper.
git-repo: https://github.com/andrewheiss/who-cares-about-crackdown
abstract: >-
  Over the past two decades, international non-governmental organizations (INGOs) have faced increasing legal scrutiny and repression abroad. As a result, official aid to INGOs in repressive countries has reduced. NGOs thus need access to a diverse array of funds and may need to increase fundraising from private donors. However, research on the determinants of philanthropy towards international causes—particularly towards organizations working in challenging environments—is still in its infancy. Which demographic and experiential factors influence the decisions of individuals to donate to legally besieged NGOs? We investigate this question using a market simulation based on a nationally representative conjoint experiment conducted on individuals across the U.S. with a history of charitable giving. We find that trust in political institutions and charities is enormously influential in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to human rights INGOs with friendly relationships with host governments, and that support holds steady if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do not face criticism or crackdown abroad. Our findings have practical implications for INGOs raising funds from individuals amid closing civic space.
reference-section-title: References
toc: false
title-page: true
word-count: "7,562"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression, international giving, donor choice
thanks: >-
  The data that support the findings of this study, as well as the code to reproduce the findings, are openly available at <https://github.com/andrewheiss/who-cares-about-crackdown> and <https://osf.io/sm5ew/>. This experiment received approval by the human subjects research committees at Christopher Newport University (1436622-1), Brigham Young University (E19135), and Georgia State University (H19644). As part of a larger project, we preregistered hypotheses and our research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/hsbyd/>. This study covers the third research question of our preregistered design.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Department of International Affairs, Lewis \& Clark College; Marc Dotson, PhD, Assistant Professor of Marketing, Marriott School of Business, Brigham Young University; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
correspondence: >-
  Correspondence concerning this article should be addressed to Andrew Heiss, PhD, Department of Public Management and Policy, Andrew Young School of Policy Studies, Georgia State University, 14 Marietta Street NW, Suite 357, Atlanta, GA 30303. Contact: aheiss@gsu.edu
mainfont: Lora
sansfont: IBM Plex Sans
policy-implications: >-
  In the wake of the global crackdown on civil society, international NGOs (INGOs) need to maintain access to multiple sources of funding, including individual donors.

@@

In 2016, Human Rights Watch claimed that civil society was under more aggressive attack than at any time in recent memory [@Roth:2016]. Governments across the globe have cracked down on civil society organizations (CSOs) and non-governmental organizations (NGOs)—a phenomenon known as "closing civic space" [@CarothersBrechenmacher:2014; @CIVICUS:2017; @DPR:2015; @DPR:2016].[^NGOdef] As part of this phenomenon, NGOs face strong legal and bureaucratic obstacles to their work. Legal restrictions, or what we refer to as legal crackdowns, are a core part of these efforts. These crackdowns create barriers to entry, funding, and advocacy for NGOs in an effort to control, obstruct, and repress these organizations. Barriers to funding are especially pervasive and restrict the ability of NGOs to secure financial resources. Restrictive states may prevent the transfer of foreign funds to NGOs based on the origin of these funds and which issues these funds can be used towards.

As a result of these laws, foreign aid (also referred to as public or official aid) channeled through NGOs has decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. Aid from public organizations like the United States Agency for International Development (USAID) and the Organisation for Economic Cooperation and Development (OECD) has decreased following the passage and implementation of these laws. However, philanthropy from private donors and foundations is not as adversely affected [@McGill:2018]. Many foundations have continued channeling funds to countries with hostile legal environments [@NeedlesWolcheckSato:2018]. Given this trend of reduced public aid following restricted legal environments for NGOs abroad, many INGOs may need to devote more resources to fundraise from individual donors.

Individual private donations are a critical source of INGO funding. While total private philanthropy[^phildef] in the United States has grown substantially over time, particularly notable is the fact that private giving to NGOs working in international affairs  and international affairs-focused organizations are one of only two categories of organizations that have continued experiencing substantial growth in giving and reached \$22.08 billion in 2018 [@givingusa:2019].[^intlaffairs] Most of this growth is driven by individuals making small-scale or micro-donations.[^smalldonors] In 2017, 70% of donations to international affairs NGOs came from individuals, rising to 86% in 2018 [@givingusa:2019]. But relying on individual donors means that NGOs must consider their preferences when fundraising. Which individual donor characteristics and attributes are most salient for INGOs interested in maximizing philanthropy at a time when they are facing increasingly hostile environments abroad? What kinds of donors consider an INGO's host country political environment when making the decision to donate?

@@

We explore this question using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity. We adapt methods used in marketing to generate dozens of distinct simulated personas that represent typical donor profiles, and we vary these personas along different dimensions of demographic, political, and social attributes. We find that specific individual characteristics help shape donor preferences for donating to INGOs, and that an organization's target issue area and its relationship with its host government interact strongly with these characteristics. Personas with high levels of trust in political and charitable institutions and who are actively engaged in charity and volunteerism prefer donating to human rights organizations that have friendly relationships with their host governments. In contrast, personas with low levels of social trust eschew both human rights organizations *and* any type of NGO with more contentious relationships with host governments. Instead, they prefer donating to NGOs with friendly host-country relationships and working on less contentious issues like emergency response and refugee relief. Personas with higher trust are more comfortable donating to legally besieged INGOs. Our findings imply that individual donors with high trust would continue remaining supportive of INGOs when they face criticism and crackdown from foreign governments. On the other hand, government crackdown, and more broadly, closing civic space, may further sour NGOs least trusting individual donors.

This study contributes both to scholarship and practice in public policy, nonprofits, and philanthropy in three important ways. The paper helps us understand how the international community and NGOs might respond to the phenomenon of closing civic space, which does not appear to be receding. Most NGOs working internationally have previously relied on government and foundation funding. But in the face of decreasing funds from official donors, our results show that INGOs can rely on individual donors, specifically, their most trusting donors, when groups face a hostile environment abroad. These findings lend support to previous literature arguing that trust matters only when individuals donate to organizations acting under great uncertainty or working towards difficult goals [@Wiepking:2010].

@@

LINE ABOUT EXTENT. This phenomenon can have three main impacts on INGOs. First, it has led to a reduction in resources and grants available to INGOs. Second, restrictive NGO laws can affect INGO programming by preventing organizations from working in certain issue areas, or at least by making it more burdensome to do so. To deal with this, some groups have adapted by changing the nature of their programming to appear less threatening to the repressive countries. Finally, in the worst-case scenario, this phenomenon can also lead to a loss in INGOs access to target countries—often also associated with the use of violence against INGOs and their employees. In this paper, we exlusively focus on the first dilemma—the reduction in INGO funding and possible responses to it.

@@

Thousands of INGOs in the Global South receive funds from a variety of public and private sources each year. Public funds—or traditional donor aid—channel money from official aid agencies through a variety of bilateral and multilateral institutions. Private sources include foundations, private corporations, and individuals. While INGOs have not been passive in responding to the crackdown on civil society, they face increasing obstacles from a funding perspective.

@@

We argue that individual private donors can help fill in an important gap. Previous research has shown that individual donors do not use the same performance-based metrics as official donors [@desaikharas]. Rather, individual donors often use heuristics to simply their decision-making, making cursory judgments about an organization’s issue area, mission, vision, and values and seek out supplementary information from friends family, and acquaintances [@CrosonShang:2011a; @TBPrakash:2017; @SzperPrakash:2011]. Due to these differences, we argue that individual donors can help address some of the challenges arising from reduced public funding available to INGOs. To be clear, we are not arguing that giving by individual donors would be a perfect substitute for NGOs net loss of income or income for certain programs, but rather, an important supplement to reduced funding from foreign sources. Moreover, it would also be a supplement to private funding from foundations towards international nonprofits which has not been as adversely affected, with foundations continue to channel funds to countries with hostile legal environments [@NeedlesWolcheckSato:2018].

@@

 the role of higher education, levels of religiosity, and trust.

@@

We explore this question using a conjoint experiment conducted with a nationally representative sample in the U.S. Importantly, because there is so little research about international donations during legal crackdown by host governments, our analysis and results are exploratory and descriptive. Accordingly, we do not test specific hypotheses or theories. However, we do talk about causal effects given that we conduct an experiment that varies a number of factors in order to elicit opinions from respondents. Instead, we structure our analysis around broader themes from the existing literature and investigate how donors' (1) demographic characteristics, (2) public affairs knowledge and experience, and (3) levels of trust in government, charitable work, and associational life change individual donation preferences across different organizational issue areas, relationships with host governments, and funding sources. Our innovative experimental approach provides rich quantitative details of individual and organizational characteristics simultaneously and reveals unique insights into donor motivations that both confirm and expand previous research.

@@

In general, regardless of income, personas respond better to human rights and refugee assistance organizations, followed closely by emergency response organizations. Environmental NGOs have the lowest market share across both levels of incomes. There are negligible differences across income levels—wealthier personas show a slight preference towards human rights and refugee organizations. As we explore below, this is possibly because income correlates with interest in public affairs and political activity—individuals who are wealthier arguably have more time and resources to follow the news and be continuously engaged in their communities and in other activist causes. There are no sizable differences in donation preferences by income across different forms of NGO funding—both levels of income tend to prefer organizations funded by government grants over small donors.

An organization's relationship with its host government does influence market shares across levels of income, but the size of this difference depends on the magnitude of income. When splitting income based on income, there is no difference between income levels. In general, personas prefer to donate to organizations with friendly relationships with their host governments and avoid those that are criticized or under crackdown. However, instead of looking at personas with below- and above-median incomes, we can simulate a subset of wealthier donors. These personas are more likely to prefer donating to organizations that are criticized by their host governments. This falls in line with their preference for human rights organizations, which have more contentious missions and are more likely to receive government criticism. This support, however, appears to evaporate as civil society crackdown intensifies. High-income personas are the least likely demographic to prefer donating to NGOs under crackdown. Their lower income counterparts, on the other hand, are generally unfazed by NGOs' relationships with government, maintaining the same average market share across all three types of host country relationships. Patterns in preferences for issue area and funding sources remain constant across both methods of categorizing income—crackdown only has an effect on very high-income personas. Income thus appears to influence preferences particularly for NGOs facing criticism and crackdown abroad—wealthy donors are perhaps fair-weather friends that are more likely to pull their support when an organization's work becomes more difficult.

@@

While education and religiosity are both important drivers of individual philanthropy, neither are associated with any sizable differences in persona donation patterns. As seen in @fig:edu-religion, preferences for issue areas are constant across both persona characteristics, with human rights and refugee relief organizations the most preferred, followed closely by emergency response organizations, regardless of levels of education or religiosity. Human rights issues attract a slightly larger market share of personas who are college graduates or who attend religious services at least monthly, while refugee issues are more likely to attract less religious personas, but these differences are minor. Preferences are consistent for different sources of funding as well—personas generally prefer organizations funded mostly by government grants. Persona preferences across different organizational relationships with host governments are also consistent across education and religiosity, with most personas preferring to donate to NGOs that enjoy friendly relations abroad. College graduate personas are slightly more willing to donate to organizations facing crackdown, but the difference is roughly half a percentage point.

@@

The global crackdown on civil society and the consequent decline of official development assistance has made the role of philanthropy in sustaining INGOs clear. INGOs need access to multiple sources of funding, and the value of individual-level philanthropy in sustaining the operations of groups working in challenging environments is still relatively under-explored. While trust is a difficult construct to define, the nonprofit literature conceptualizes trust as a key driver of commitment, which is shown to impact both intention to donate as well as actual donation behavior. When considering repressive environments and NGOs working on more difficult goals, a lack of trust could be further fueled by legal trouble with the NGO host government—a dynamic that individuals may not wish to enter with their dollars. How does trust in NGOs influence individuals' preferences to donate, especially when groups face crackdown?

Using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity, we find several important trends. Trust in political institutions and in charitable organizations matters substantially in shaping donor preferences. Personas with high levels of social and charitable trust prefer donating to human rights organizations that have friendly relationships with their host governments, while personas with low trust are drawn more to emergency response and refugee relief organizations with unfriendly relationships with their host governments. Personas with low social trust tend to be wary of negative host-country relationships and strongly prefer to donate to organizations that are friendly with their host governments, while personas with high social trust are more willing to stick with legally besieged organizations. This could imply who donors who are the most trusting and involved could remain supportive when charities face government criticism and crackdown. The global crackdown on NGOs may also possibly sour NGOs' least trusting individual donors. Our findings lend support to research showing that individuals who trust NGOs are more likely to continue their monetary support of NGOs in the face of difficulties and under great uncertainty [@dwyer1987; @Wiepking:2010].

In addition, we find that on average, human rights organizations tend to attract personas who are wealthier, follow national and international news, and travel abroad. Donors who are less wealthy, follow the news less often, and do not travel, are more attracted to organizations working on less contentious like emergency response and refugee relief.

The above results have important implications for NGOs as they navigate a shrinking civic space and consider the use of different frames to tailor their fundraising appeals to individual donors. Our simulation finds that at least for some kinds of personas, NGOs may benefit from publicizing when they are targets of government crackdown, including that information in their fundraising appeals. However, the larger challenge may be addressing the concerns of donors with low levels of trust, as they are the least likely to donate organizations facing criticism and crackdown abroad. More broadly, studies from a range of countries show that levels of giving are lowest for refugee and asylum organizations, as well as for overseas and development organizations. Rather, donors giving to international causes tend to support international disaster relief and religious groups [@RobsonHart:2020]. However, some of the former are precisely the organizations facing crackdown. More research is needed to answer how these organizations can appeal to broader array of potential foreign donors to increase levels of charitable-giving towards them.

@@

[^NGOdef]: Broadly speaking, NGOs are any local, national, or international not-for-profit, voluntary organization. They can be domestic (i.e. an NGO that operates entirely in a single state), or international (INGOs). INGOs are composed of members from two or more countries and are organized to advance their members' international goals and provide services to citizens of other states through routine transactions with states, private actors, and international institutions [@Tarrow:2001, 12].

@@

[^phildef]: We use philanthropy to connote voluntary giving by an individual or group to promote the common good. While it can include grants given by foundations and corporations to nonprofit organizations, our focus here is on individual giving. Some perspectives see philanthropy as more of a strategic long-term practice that seeks to address the root causes of issues, and charity as a more short-term practice focused on providing immediate relief. But many scholars see philanthropy as essentially synonymous with charitable donations, as the latter can also address longer-term or strategic isssues [@Castle:2004]. As such, despite these subtle differences, we use the terms inter-changeably here both because previous literature does so, and because in the case of philanthropic or charitable-giving to NGOs during crackdown, individuals could be giving to nonprofits both for immediate relief as well as long-term strategic purposes.

>>>>

---
title: "Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy"
short-title: "Who Cares About Crackdowns?"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
date: "October 2, 2020"
published: Working paper.
git-repo: https://github.com/andrewheiss/who-cares-about-crackdown
abstract: >-
  The phenomenon of closing civic space has adversely impacted the operations of INGOs. It has not just resulted in net losses of income to INGOs, but also how and where funds are allocated. We argue that individual private donors can be important in sustaining the operations of INGOs working in repressive contexts. Individual donors do not use the same performance-based metrics as official aid donors, and trust has shown to be an important component of support for nonprofits working towards difficult goals. How does trust in charitable organizations influence individuals’ preferences to donate, especially when these groups face crackdown? Using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity, we find that trust in charitable organizations matters substantially in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to INGOs with friendly relationships with host governments. This support holds steady if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do not face criticism or crackdown abroad. The global crackdown on NGOs may thus possibly sour NGOs' least trusting individual donors. Our findings have practical implications for INGOs raising funds from individuals amid closing civic space.
reference-section-title: References
toc: false
title-page: true
word-count: "7,562"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression, international giving, donor choice
thanks: >-
  The data that support the findings of this study, as well as the code to reproduce the findings, are openly available at <https://github.com/andrewheiss/who-cares-about-crackdown> and <https://osf.io/sm5ew/>. This experiment received approval by the human subjects research committees at Christopher Newport University (1436622-1), Brigham Young University (E19135), and Georgia State University (H19644). As part of a larger project, we preregistered hypotheses and our research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/hsbyd/>. This study covers the third research question of our preregistered design.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Department of International Affairs, Lewis \& Clark College; Marc Dotson, PhD, Assistant Professor of Marketing, Marriott School of Business, Brigham Young University; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
correspondence: >-
  Correspondence concerning this article should be addressed to Andrew Heiss, PhD, Department of Public Management and Policy, Andrew Young School of Policy Studies, Georgia State University, 14 Marietta Street NW, Suite 357, Atlanta, GA 30303. Contact: aheiss@gsu.edu
mainfont: Lora
sansfont: IBM Plex Sans
policy-implications: >-
  In the wake of the global crackdown on civil society, international NGOs (INGOs) need to maintain access to multiple sources of funding, including individual donors.

@@

In 2016, Human Rights Watch claimed that civil society was under more aggressive attack than at any time in recent memory [@Roth:2016]. Governments across the globe have cracked down on civil society organizations (CSOs) and non-governmental organizations (NGOs)—a phenomenon known as "closing civic space" [@CarothersBrechenmacher:2014; @CIVICUS:2017; @DPR:2015; @DPR:2016].[^NGOdef] Legal restrictions, or what we refer to as legal crackdowns, are a core part of these efforts. These crackdowns create barriers to entry, funding, and advocacy for NGOs in an effort to control, obstruct, and repress these organizations. Barriers to funding are especially pervasive and restrict the ability of NGOs to secure financial resources. Restrictive states may prevent the transfer of foreign funds to NGOs based on the origin of these funds and which issues these funds can be used towards.

As a result of these laws, foreign aid (also referred to as public or official aid) channeled through NGOs has decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. However, philanthropy from private donors and foundations is not as adversely affected [@McGill:2018]. Many foundations have continued channeling funds to countries with hostile legal environments [@NeedlesWolcheckSato:2018]. Given this trend of reduced public aid following restricted legal environments for NGOs abroad, INGOs may need to devote more resources to fundraise from individual donors.

Individual private donations are a critical source of INGO funding. While total private philanthropy[^phildef] in the United States has grown substantially over time, it is notable that private giving to NGOs working in international affairs  and international affairs-focused organizations are one of only two categories of organizations that have continued experiencing substantial growth in giving and reached \$22.08 billion in 2018 [@givingusa:2019].[^intlaffairs] Most of this growth is driven by individuals making small-scale or micro-donations.[^smalldonors] In 2017, 70% of donations to international affairs NGOs came from individuals, rising to 86% in 2018 [@givingusa:2019]. But relying on individual donors means that NGOs must consider their preferences when fundraising. Which individual donor characteristics and attributes are most salient for INGOs interested in maximizing philanthropy at a time when they are facing increasingly hostile environments abroad?

@@

We explore this using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity. We adapt methods used in marketing to generate dozens of distinct simulated personas that represent typical donor profiles, and we vary these personas along different dimensions of demographic, political, and social attributes. We find that an organization's target issue area and its relationship with its host government interact strongly with individuals' characteristics. Personas with high levels of trust in political and charitable institutions and who are actively engaged in charity and volunteerism prefer donating to human rights organizations that have friendly relationships with their host governments. In contrast, personas with low levels of social trust eschew both human rights organizations *and* any type of NGO with more contentious relationships with host governments. Instead, they prefer donating to NGOs with friendly host-country relationships and working on less contentious issues like emergency response and refugee relief. Personas with higher trust are more comfortable donating to legally besieged INGOs. Our findings imply that individual donors with high trust would continue remaining supportive of INGOs when they face criticism and crackdown from foreign governments. On the other hand, government crackdown may further sour NGOs least trusting individual donors.

This study contributes both to scholarship and practice in public policy, nonprofits, and philanthropy in three important ways. It helps us understand how the international community and NGOs might respond to the phenomenon of closing civic space, which does not appear to be receding. Most NGOs working internationally have previously relied on government and foundation funding. But in the face of decreasing funds from official donors, our results show that INGOs can rely on individual donors, specifically, their most trusting donors, when groups face a hostile environment abroad. These findings lend support to previous literature arguing that trust matters  when individuals donate to organizations acting under great uncertainty or working towards difficult goals [@Wiepking:2010].

@@

Mre than 100 countries have proposed or enacted 244 measures restricting, repressing, or shutting down civil society since 2013 (ICNL 2021 CITATION FROM ISA PAPER). This has had a three-hold impact on INGOs. First, it has led to a reduction in resources and grants available to INGOs. Second, restrictive NGO laws can affect INGO programming by preventing organizations from working in certain (contentious) issue areas. To deal with this, some groups have adapted by changing the nature of their programming to appear less threatening to the repressive countries. Finally, in the worst-case scenario, this has led to a loss in INGOs' access to target countries—often also associated with the use of violence against INGOs and their employees. In this paper, we exlusively focus on the first dilemma—the reduction in INGO funding and possible responses to it.

@@

Many INGOs in the Global South receive funds from a variety of public and private sources each year. Public funds—or traditional donor aid—channel money from official aid agencies through a variety of bilateral and multilateral institutions. Private sources include foundations, private corporations, and individuals. While INGOs have not been passive in responding to the crackdown on civil society, they face increasing obstacles from a funding perspective.

@@

We argue that individual private donors can help fill in an important gap. Individual donors do not use the same performance-based metrics as official donors [@desaikharas]. Rather, individual donors often use heuristics to simply their decision-making, making cursory judgments about an organization’s issue area, mission, vision, and values and seek out supplementary information from friends, family, and acquaintances [@TBPrakash:2017; @SzperPrakash:2011]. Due to these differences, individual donors can help address some of the challenges arising from reduced public funding available to INGOs. To be clear, we are not arguing that giving by individual donors would be a perfect substitute for NGOs net loss of income or income for certain programs, but rather, an important supplement to reduced funding from foreign sources. Moreover, it would also be a supplement to private funding from foundations towards international nonprofits which has not been as adversely affected, with foundations continue to channel funds to countries with hostile legal environments [@NeedlesWolcheckSato:2018].

@@

## Key drivers of individual-level philanthropy

Substantial research has looked at the individual-level determinants of donating to local, rather than international causes, including four main attributes: social capital, the role of higher education, levels of religiosity, and trust.

@@

We explore this question using a conjoint experiment conducted with a nationally representative sample in the U.S. Importantly, because there is so little research about international donations during legal crackdown by host governments, our analysis and results are exploratory and descriptive. Accordingly, we do not test specific hypotheses or theories.

TODO: Fix final point - connect testing stuff with existing theories - we don't have a list of things to test

However, we do talk about causal effects given that we conduct an experiment that varies a number of factors in order to elicit opinions from respondents. Instead, we structure our analysis around broader themes from the existing literature and investigate how donors' (1) demographic characteristics, (2) public affairs knowledge and experience, and (3) levels of trust in government, charitable work, and associational life change individual donation preferences across different organizational issue areas, relationships with host governments, and funding sources. Our innovative experimental approach provides rich quantitative details of individual and organizational characteristics simultaneously and reveals unique insights into donor motivations that both confirm and expand previous research.

@@

Regardless of income, personas respond better to human rights and refugee assistance organizations, followed closely by emergency response organizations. Environmental NGOs have the lowest market share across both levels of incomes. There are negligible differences across income levels—wealthier personas show a slight preference towards human rights and refugee organizations. This is possibly because income correlates with interest in public affairs and political activity—individuals who are wealthier arguably have more time and resources to follow the news and be continuously engaged in their communities and in other activist causes. There are no sizable differences in donation preferences by income across different forms of NGO funding—both levels of income tend to prefer organizations funded by government grants over small donors.

An organization's relationship with its host government does influence market shares across levels of income, but the size of this difference depends on the magnitude of income. In general, personas prefer to donate to organizations with friendly relationships with their host governments and avoid those under crackdown. Wealthier donor personas are more likely to prefer donating to organizations that are criticized by their host governments. This falls in line with their preference for human rights organizations, which have more contentious missions and are more likely to be targeted. This support, however, evaporates as crackdown intensifies. High-income personas are the least likely demographic to prefer donating to NGOs under crackdown. Their lower income counterparts, on the other hand, are generally unfazed by NGOs' relationships with government, maintaining the same average market share across all three types of host country relationships. Patterns in preferences for issue area and funding sources remain constant across both methods of categorizing income—crackdown only has an effect on very high-income personas. Income thus appears to influence preferences particularly for NGOs facing crackdown abroad—wealthy donors are perhaps fair-weather friends that are more likely to pull their support when an organization's work becomes more difficult.

@@

While education and religiosity are both important drivers of individual philanthropy, neither are associated with any sizable differences in persona donation patterns (as seen in @fig:edu-religion). Human rights issues attract a slightly larger market share of personas who are college graduates or who attend religious services at least monthly, while refugee issues are more likely to attract less religious personas, but these differences are minor. Preferences are consistent for different sources of funding as well—personas generally prefer organizations funded mostly by government grants. Persona preferences across different organizational relationships with host governments are also consistent across education and religiosity, with most personas preferring to donate to NGOs that enjoy friendly relations abroad.

@@

The global crackdown on civil society has adversely impacted the operations of NGOs. It has not just resulted in net losses of income from official sources to INGOs, it has also affected how and where funds are allocated. We argue that individual donors can be useful in sustaining the operations of groups working in challenging environments. Individual donors do not use the same perfomance-based metrics as official aid donors and use heuristics to smiplify their decision-making. The role of trust may be important in understanding donations preferences towards INGOs working in challenging contexts. While trust is a difficult construct to define, the nonprofit literature conceptualizes trust as a key driver of commitment, which is shown to impact both intention to donate as well as actual donation behavior. With repressive environments and NGOs working on more difficult goals, a lack of trust could be further fueled by legal trouble with the NGO host government—a dynamic that individuals may not wish to enter with their dollars. How does trust in NGOs influence individuals' preferences to donate, especially when groups face crackdown?

Using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity, we find several important trends. Trust in political institutions and in charitable organizations matters substantially in shaping donor preferences. Personas with high levels of social and charitable trust prefer donating to human rights organizations that have friendly relationships with their host governments, while personas with low trust are drawn more to emergency response and refugee relief organizations with unfriendly relationships with their host governments. Personas with low social trust tend to be wary of negative host-country relationships and strongly prefer to donate to organizations that are friendly with their host governments, while personas with high social trust are more willing to stick with legally besieged organizations. This implies that donors who are the most trusting could remain supportive when nonprofits face government criticism and crackdown. The global crackdown on NGOs may also possibly sour NGOs' least trusting individual donors. Our findings lend support to research showing that individuals who trust NGOs are more likely to continue their monetary support of NGOs in the face of difficulties and under great uncertainty [@dwyer1987; @Wiepking:2010].

Our results have important implications for NGOs as they navigate a shrinking civic space and consider the use of different frames to tailor their fundraising appeals to individual donors. Our simulation finds that at least for some kinds of personas, NGOs may benefit from publicizing when they are targets of government crackdown, including that information in their fundraising appeals. However, the larger challenge may be addressing the concerns of donors with low levels of trust, as they are the least likely to donate organizations facing criticism and crackdown abroad. As @TBPrakash:2017 argue, "trust in the charity sector cannot be assumed: the sector has to demonstrate that the trust it receives is justified." (643).

When considering INGOs working in different issue areas, studies from a range of countries show that levels of giving are lowest for refugee and asylum organizations. Rather, donors giving to international causes tend to support international disaster relief and religious groups [@RobsonHart:2020]. However, some of the former are precisely the organizations facing crackdown. More research is needed to answer how contentious or advocacy organizations (as opposed to service nonprofits) can appeal to broader array of potential foreign donors to increase levels of charitable-giving towards them.

@@

[^NGOdef]: Broadly speaking, NGOs are any local, national, or international not-for-profit, voluntary organization. They can be domestic (i.e. an NGO that operates entirely in a single state), or international (INGOs). INGOs are composed of members from two or more countries and are organized to advance their members' international goals and provide services to citizens of other states through transactions with states, private actors, and international institutions [@Tarrow:2001, 12].

@@

[^phildef]: We use philanthropy to connote voluntary giving to promote the common good. While it can include grants given by foundations and corporations to nonprofit organizations, our focus is on individual giving. Some perspectives see philanthropy as more of a strategic long-term practice addressing the root causes of issues, and charity as a more short-term practice focused on providing immediate relief. But many scholars see philanthropy as essentially synonymous with charitable donations, as the latter can also address longer-term or strategic isssues [@Castle:2004]. Despite these subtle differences, we use the terms inter-changeably here both because previous literature does so, and because in the case of giving to NGOs during crackdown, individuals could be giving to nonprofits both for immediate relief as well as long-term strategic purposes.
